Initial Analysis

### 2019-07-19 Initial_Analysis.Rmd
Modified code to address issues discussed with Gabriel. Now SAU catch data works as a treshold and area index is fixed

# The function
EstTransIndex <- function(Spp,Model,Neighbors,Coord){

@@

  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX) %>% # Filter data only located within EEZs
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  # For now using SAU_C as presence absence
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)
              ) %>%
    filter(Model_Index > 0) %>%
    mutate(Model_Index = Model_Index*100) %>%
    left_join(Neighbors,
              by = "INDEX")

  # Estimating total grids per country
  Spp_Grid <- Trans_Spp %>% 
    group_by(Name = Country_Territory,
             Neighbor_Territory,
             Model_Index) %>% 
    summarise(n_cells_spp = n()) %>% 
    ungroup()

  # Split dataframes to merge latter
  Tone <- select(Spp_Grid,1,3,4)
  Ttwo <- select(Spp_Grid,Name=2,3,4) # Change column name

  # Merge dataframes
  Merged_data <- full_join(Tone,Ttwo, by = c("Name","Model_Index")) %>% 
    rowwise() %>% 
    mutate(Spp_Total = sum(n_cells_spp.x,n_cells_spp.y,na.rm=T)) %>% 
    select(Name,Model_Index,Spp_Total) %>% 
    left_join(Spp_Grid,
              by = c("Name","Model_Index")) %>% 
    ungroup() %>% 
    group_by(Name,Model_Index) %>% 
    summarise(Spp_Total = sum(Spp_Total))

@@

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Indexes <- Trans_Spp %>% 
    group_by(Name,
             Neighbor_Territory,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% # The number of cells present within each country's EEZ
    #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 2)_________ #
    left_join(Merged_data,
              by = c("Name","Model_Index")) %>% 
    mutate(Trans_Index = n_cells_spp/Spp_Total)

@@

  if(nrow(Indexes) > 0){

@@

### Testing Function ####
# Spp <- 600006
# Model <- "All"
# Neighbors <- Neighbor_List

@@

@@

# TI of 10% and MI > 50

TenFifty <- Results_Data %>% 
  filter(Model_Index >= 50#,
         # Trans_Index >= 0.10
         ) %>% 
  group_by(Name,Model_Index) %>% 
  summarise(n_Trans_Spp = n()) %>% 
  filter(Model_Index == 100)
  # head(Results_Data)

@@

>>>>

# The function
EstTransIndex <- function(Spp,Model,Neighbors,Coord,Save="Y"){

@@

@@

  #____________ ESTIMATING MODEL INDEX (TRESHOLD 1)_________ #
  Trans_Spp <- SppDist %>%
    filter(INDEX %in% Neighbors$INDEX,# Filter data only located within EEZs
           Model != "SAU_C") %>% # Only using observational and modelled data
    mutate(Value = ifelse(Value > 0, 1,0)) %>%  
    group_by(TaxonKey,
             INDEX
    ) %>%
    summarise(Model_Index = mean(Value,na.rm=T)
              ) %>%
    filter(Model_Index > 0) %>%
    #____________ ESTIMATING FUNDAMENTAL NICHE (TRESHOLD 2)_________ #
    left_join(SppDist,
              by = c("TaxonKey","INDEX")) %>% 
    filter(Model == "SAU_C", # Only keeping cells where SAU catch exists
           Value > 0) %>% 
    mutate(Model_Index = Model_Index*100) %>% 
    select(-Model, -Value, -Latitude,-Longitude) %>% 
    left_join(EEZ_CellID,
              by = "INDEX") %>% 
    left_join(EEZIDs_List,
              by = "EEZID")

  # MODEL INDEX (TRESHOLDS 1 & 2) #
  Model_Index_D <- Trans_Spp %>% 
    group_by(Name,
             TaxonKey,
             Model_Index
    ) %>% 
    summarise(n_cells_spp = n()) %>% 
    select(-n_cells_spp)

    #____________ ESTIMATING DISTRIBUTION INDEX (TRESHOLD 3)_________ #
   # The number of species' cells present within each country's EEZ

  #Step 1.  Get EEZ id and neighbor
  Neighbors_EEZ <- Neighbors %>% 
  group_by(EEZID,Country_Territory,Neighbor_Territory) %>% 
  summarise(n=n()) %>% 
  ungroup() %>% 
  select(-n)

  # Step 2. Determines the amount of grids present in each country
  Spp_Grid <- Trans_Spp %>% 
    group_by(TaxonKey,Model_Index,EEZID) %>% 
    summarise(n_spp_eez = length(unique(INDEX))) %>% 
    left_join(Neighbors_EEZ,
              by = "EEZID") %>% 
    filter(Country_Territory %in% Trans_Spp$Name, #Filter out unwanted neighbors (those who don't have grids within but get included because they are neighbors)
           Neighbor_Territory %in% Trans_Spp$Name)

  # Step 3. Sum total grids per neighbors

  # Split dataframes to merge latter
  Country_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(Model_Index,TaxonKey,Name=Country_Territory,n_spp_eez)

  Neighbor_Territory_T <- Spp_Grid %>% 
    ungroup() %>% 
    select(Model_Index,TaxonKey,n_spp_eez,Name=Neighbor_Territory, Country_Territory)

  # Merge dataframes to get totals per neighbords
  Area_Index_D <- full_join(Country_Territory_T,
                           Neighbor_Territory_T, by = c("Model_Index","Name","TaxonKey")) %>% 
    rowwise() %>%
    mutate(Spp_Total = sum(n_spp_eez.x,n_spp_eez.y,na.rm=T)) %>% # Total gridcelles per neighbors
    distinct() %>% # Removes false duplicates from `full_join()`
    rename(Neighbor_Territory = Name,
           n_spp_Country = n_spp_eez.x,
           n_spp_Neighbor = n_spp_eez.y) %>% 
    mutate(Area_Index = n_spp_Country/Spp_Total) %>%  #Estimates the proportion of grids per country
    select(1:3,6,8)

  #____________ FINAL INDEX TABLE_________ #
  Indexes <- Model_Index_D %>% 
  rename(Country_Territory = Name) %>% 
  full_join(Area_Index_D,
            by = c("TaxonKey","Country_Territory","Model_Index")
  ) %>% 
    filter(!is.na(Area_Index),
           !is.na(Model_Index)
           )%>%  # Remove cases where the country does not pass the area_index and viceversa
    select(TaxonKey,Country_Territory,Neighbor_Territory,everything())

@@

  if(nrow(Indexes) > 0 & Save == "Y"){

@@

### Testing Function ####
# Spp <- 600006
# Model <- "All"
# Neighbors <- Neighbor_List
# 
# Tryout <- EstTransIndex(Spp,Model,Neighbors,CoorG,Save="N")
# 
# Over_20 <-Tryout %>% 
#     filter(Area_Index > 0.20,
#            Area_Index < 0.80)

@@

# SAU relations between INDEX and Country's EEZs
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))
EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

@@

### Problem solving

#### Overcounting transboundary stocks

In a first run, Australia results in many transboundary stocks, I believe this is because it is counting the number of countries that that the species share (e.g. one species 6 countries = 6 transboundary spp) rather than 1 speces 6 countries = 1 trans spp.

**Problem fixed.** Now we can estimate the number of transboundary species as well as the number of countries/regions that each country shares per species.

```{Australia_case, eval =F, echo = F}

Australia <- Results_Data %>% 
  filter(Name == "Australia")

Original_Australia <-Australia %>% 
  filter(Model_Index >= 50#,
         # Trans_Index >= 0.10
         ) %>% 
  group_by(Name,Model_Index) %>% 
  summarise(n_Trans_Spp = n()) %>% 
  filter(Model_Index == 100)

## Testing the waters

Explore_Australia <- Australia %>% 
  filter(Model_Index >= 50#,
         # Trans_Index >= 0.10
         ) %>% 
  slice(1:20) %>% 
  group_by(Name,Model_Index) %>% 
  summarise(
    n_trans = length(unique(TaxonKey)),
    n_countries = length(unique(Neighbor_Territory)),
    )

#### Fixing counts within EEZs

In the first run. Countries where EEZs are devided but touching (e.g. South Africa inidan and atlantic coasts) are double counted.

This might be fixed with Tayler's data of EEZs to ISOs

```{South_Africa_case, eval =F, echo = F}

# Tyler E.'s' data
EEZ_ISO <- read.csv("/Volumes/DATA/JULIANO_NEYMAR/FishForVisa/Spatial_Data/EEZID_ISO_TylEd.csv", comment.char="#")

#SAU / Vicky's data
EEZIDs_List <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/Updated_EEZList_17June2016.xlsx",sep=""))

EEZ_CellID <- read_excel(paste(Data_Path,"Spatial_Data/V_Lam/EEZ_CellID.xlsx",sep=""))
colnames(EEZ_CellID) <- c("EEZID","INDEX")

# Worlds adata

# The path
path_world <- paste(Data_Path,"Spatial_Data/World_EEZ_v10_2018",sep="")
# #The File
fnam_world <- "eez_v10.shp"
# #Load it!
# 
World_EEZs <- st_read(dsn = path_world,
                      layer =file_path_sans_ext(fnam_world))

head(World_EEZs)

World_EEZs %>% 
  filter(
    # Territory1 == "South Africa",
    ISO_Ter1 == "ZAF") %>% 
  head()

# Get ISO and territory names from world map
ISO_Ter <- World_EEZs %>%
  select(ISO_Ter1,Territory1) %>% 
  as.data.frame() %>% 
  select(ISO_Ter1,Territory1)

head(ISO_Ter)

# I think EEZID equals ISO when not devided...

EZZID_Yes_ISO <- ISO_Ter %>% 
  semi_join(ISO_Ter,
            by ="ISO_Ter1") %>% 
  rename(Name = Territory1) %>% 
  # head() %>% 
  left_join(EEZIDs_List,
            by = "Name")

EEZ_ISO_Clean <- ISO_Ter %>% 
  left_join(EEZ_ISO,
            by = "ISO_Ter1")

SA <- Results_Data %>% 
  filter(stringr::str_detect(Name,"South Africa"))

```{Number_Tran_country, eval =T, echo =F}
# TI of 10% and MI > 50

TenFifty <- Results_Data %>% 
  filter(Model_Index >= 50#,
         # Trans_Index >= 0.10
         ) %>% 
  group_by(Name,Model_Index) %>% 
  summarise(
    n_trans = length(unique(TaxonKey)),
    n_countries = length(unique(Neighbor_Territory))
    ) 
  filter(Model_Index == 100)
  # head(Results_Data)

@@

  Area_Index_D <- Merged_data %>% 
    filter(Area_Index > 0.10,
           Area_Index < 0.90)
